---
title: "Amsterdam Airbnb Price Prediction"
author: "Nawal Zehra Hasan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty: default
  html_document: rmdformats::robobook
---


```{r libraries, include = FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
# Import libraries
library(rmdformats)
library(stargazer)
library(kableExtra)
#install.packages("ranger")
library(ranger)
library(tidyverse)
#install.packages("rmdformats")
library(rmdformats)
library(stringr) 
library(rstatix)
library(scales)
#install.packages("Metrics")
library(Metrics)
library(caret)
library(fastDummies)
library(ggpubr)
library(cowplot)
library(dplyr)
library(skimr)
library(modelsummary)
library(grid)
#install.packages("glmnet")
library(glmnet)

```

```{r Import data, message=FALSE, warning=FALSE, include=FALSE, cache=TRUE}
# Import data
data <- read_csv("data.csv")
```
# EXECUTIVE SUMMARY

The idea of the following report is to provide statistical models capable of predicting the short rental prices in Amsterdam of apartments accommodating between 2 to 6 persons. With a project like these Airbnbs rentals can estimate at what price should they rent out their apartments and what the profit margins will be at that given price. Using 3 machine learning algorithms; OLS Linear regression, lasso, and Random Forest, I was able to perform this analysis. 

## THE DATASET

To accomplish the given task, I have used the Airbnb data for Amsterdam for the date 7th September 2021 that is available on [Inside Airbnb](www.insideairbnb.com) website. The dataset has a single table that includes 16116 observations.

The target variable given is price per night in Euros. The predictors consist of several variables related to property reviews, neighborhood, amenities offered etc. The data set had several variables that would reflect the quality of an Airbnb rental. This shows that data collection was thorough and detailed.

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- FIRST LOOK INTO DATA ------------------------ #
glimpse(data)
skim(data)

# Filtering for apartments which accommodate 2-6 persons
data <- data %>%  filter(between(accommodates, 2, 6))

# summary statistics for the data
data %>%
  group_by(accommodates) %>%
  dplyr::summarize(mean_price = mean(price, na.rm=TRUE))
Hmisc::describe(data$price)

# Check to see if there are any duplicates
duplicated(names(data))

# Drop unused columns
data <- data[grep("^host", colnames(data), invert = TRUE)]
data <- data %>% dplyr::select(-contains("maximum"))
data <- data[grep("^calculated", colnames(data), invert = TRUE)]
data <- data %>% select(-c("listing_url","scrape_id","last_scraped","name","description","neighborhood_overview","picture_url", "neighbourhood_group_cleansed","bathrooms","minimum_minimum_nights","minimum_nights_avg_ntm","calendar_updated","calendar_last_scraped","number_of_reviews_ltm","number_of_reviews_l30d","license","reviews_per_month","availability_30","availability_60","availability_90","availability_365","neighbourhood","has_availability", "first_review", "last_review"))

## Create Numerical variables
data <- data %>%
  mutate(
    price_day = as.numeric(gsub("[^0-9.]", "", price)))
```
## DATA MUNGING

The dataset required a lot of cleaning to make it suitable for use specifically for my analysis. This method of tweaking the data by filtering variables, renaming variables and dealing with missing values etc is the first and most important part of the analytical process. Our results depend on it so this took me quite some time. 

I deleted some unused columns and renamed some columns for easier use in the modeling process. I changed the price variables to exclude the dollar symbol from the price making it into a numerical column. Then I dropped all observations that had price missing. For many missing values in a column such as review_scores_value & review_scores_communication, I dropped these columns. I also checked for duplicates and found none. I filtered the data on property types specifically targeting apartments as a requirement by the study and changed it to a factor variables. After this we were left with 10085 observations. These property types were renamed to make it easier for use. The amenities column consisted of a vector of different amenities (like "tv, "stove", "pets") which needed to be split apart into individual binary columns. This resulted in more than 500 new columns being added. Using a function which searches for a particular key word (like "wifi" or "parking") through all column names and aggregates them into one general binary column. I chose a few amenities and dropped the rest. Lastly, I created dummy variables for all the categorical variables in our data such as amenities and availability.


```{r, include = FALSE, message=FALSE, warning=FALSE }
# --------------------------- DEAL WITH MISSING VALUES ------------------------ #

# A) drop if no target value
data <- data %>%
  drop_na(price)

# B) Impute if few
data <- data %>%
  mutate(
    beds = ifelse(is.na(beds), accommodates, beds), #assume beds=accomodates
    minimum_nights=ifelse(is.na(minimum_nights),1, minimum_nights),
    number_of_reviews=ifelse(is.na(number_of_reviews),1, number_of_reviews),
    beds=ifelse(is.na(beds),0, beds),
  )

# C) drop columns when many missing: review_scores_value & review_scores_communication
to_drop <- c("review_scores_value", "review_scores_communication")
data <- data %>%
  select(-one_of(to_drop))

to_filter <- sapply(data, function(x) sum(is.na(x)))
to_filter[to_filter > 0]

# D)  Replace missing variables re-reviews with zero, when no review + add flags
data <- data %>%
  mutate(
    flag_review_scores_rating = ifelse(is.na(review_scores_rating),1, 0),
    review_scores_rating    = ifelse(is.na(review_scores_rating),median(review_scores_rating, na.rm = T), review_scores_rating))

# check one flagged variable
datasummary( factor(flag_review_scores_rating) ~ N , data) 
```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- CLEAN PROPERTY TYPES ------------------------ #
# Keep if property type is Entire home/apartment, Entire loft, Entire serviced apartment
data <- data %>% 
  filter( property_type %in% c('Entire home/apt', 'Entire loft', 'Entire serviced apartment',  'Entire rental unit') )

# rename property_type categories to make them shorter
data <- data %>% mutate( property_type = 
                        ifelse( property_type == 'Entire home/apt', 'apartment', 
                        ifelse( property_type == 'Entire loft', 'loft',
                        ifelse( property_type == 'Entire serviced apartment', 'serviced_apartment',
                        ifelse( property_type == 'Entire rental unit', 'rental unit',".")))))

# convert property_type to factor
data <- data %>%
  mutate(f_property_type = factor(property_type))

# convert neighborhood_cleansed to factor
data <- data %>% 
  mutate( f_neighbourhood = factor(neighbourhood_cleansed))

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT NUMBER OF BATHROOMS ------------------------ #

data <- data  %>% 
  mutate(bathrooms = as.numeric(gsub("[a-zA-Z ]", "", data$bathrooms_text)))
data$bathrooms_text <- NULL

# add new numeric columns from certain columns
numericals <- c("accommodates", "bathrooms" , "minimum_nights","beds" ,"review_scores_rating","number_of_reviews")

data <- data %>%
  mutate_at(vars(numericals), funs("n"=as.numeric))

# rename columns so they start with n_ as opposed to end with _n
nnames <- data %>%
  select(ends_with("_n")) %>%
  names()
nnames_i <- match(nnames, colnames(data))
colnames(data)[nnames_i] <- paste0("n_", numericals)
```

```{r,include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT AMENITIES ------------------------ #

# remove unnecessary signs and convert to list
data$amenities <- tolower( data$amenities )
data$amenities <- gsub("\\[","", data$amenities)
data$amenities <- gsub("\\]","", data$amenities)
data$amenities <- gsub('\\"',"",data$amenities)
data$amenities <- as.list(strsplit(data$amenities, ","))

# define levels and dummies and append to df
levs <- levels(factor(unlist(data$amenities)))
data <- cbind(data,as.data.frame(do.call(rbind, lapply(lapply(data$amenities, factor, levs), table))))

# function to aggregate several columns of same type/category into one generic binary column
aggregate_columns <- function(word){
  
  # subset columns which contain a specific word and save them to another dataframe, also select 'id' to use for merge later
  new_df <- data %>% select(contains(word),"id")
  
  # go row by row to see if any of the rows have a 1, if it does, populate new column 'col_name' with 1
  new_df$col_name <- apply(new_df[0:ncol(new_df)], 1, function(x) ifelse(any(x == 1), '1', '0'))
  
  # save new column and id column to another dataframe, this new dataframe is used to merge with original dataframe
  new_df_merge <- new_df %>% select(id,col_name)
  
  # merge original dataframe and new_df_merge by 'id'
  data <- merge(data,new_df_merge,by = "id", all = FALSE)
  
  # remove the new column and 'id' column from the new_df dataframe
  new_df <- new_df %>% select(-c(id,col_name))

  # remove the selected columns from original dataframe since they have already been aggregated into a new column and merged
  data <<- data %>% select(-colnames(new_df))
}


# aggregate columns for a few amenities that could be important for predicting price
aggregate_columns("wifi")
data <- data %>% rename("wifi" = col_name)

aggregate_columns("refrigerator")
data <- data %>% rename("refrigerator" = col_name)

aggregate_columns("air conditioning")
data <- data %>% rename("air_conditioning" = col_name)

aggregate_columns("baby")
data <- data %>% rename("baby" = col_name)

aggregate_columns("beach")
data <- data %>% rename("beach" = col_name)

aggregate_columns("stove")
data <- data %>% rename("stove" = col_name)

aggregate_columns("free parking")
data <- data %>% rename("free_parking" = col_name)


# drop the amenities column because a csv cannot store it since it is a list
data <- data %>% select( -amenities )

# drop amenities that were not used
data <- data[ -c(30:418)]

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- CREATE DUMMY VARIABLES ------------------------ #

# create dummies
data <- data %>% mutate(
  d_instant_bookable = ifelse( instant_bookable == 't', 1, 0)
)

# rename amenities dummies
dummies <- c( "wifi", "refrigerator", "air_conditioning", "baby", "beach", "stove", "free_parking")
data <- data %>%
  mutate_at(vars(dummies), funs("d"= (.)))
# rename columns
dnames <- data %>%
  select(ends_with("_d")) %>%
  names()
dnames_i <- match(dnames, colnames(data))
colnames(data)[dnames_i] <- paste0("d_", tolower(gsub("[^[:alnum:]_]", "",dummies)))

# check if price is missing
nrow(data %>% filter( is.na(price_day)))
```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# keep columns if contain d_, n_,f_, p_, usd_ and some others
data <- data %>%
  select(id,price_day,matches("^d_.*|^n_.*|^f_.*"))
amenities_convert<- data %>%
  select(starts_with("d_"),"id") 
amenities_convert <- amenities_convert %>%mutate_if(is.integer,as.numeric)
glimpse(amenities_convert)
data <- data %>%
  select(-starts_with("d_")) 
data <- merge(data,amenities_convert, by = "id")
data <- data %>% mutate(id = as.numeric(id))
```
## EXPLORATORY DATA ANALYSIS

After the data was cleaned, I conducted Exploratory Data Analysis to try to understand the characteristics of our data. As price is the target value, it`s important to visualize it's distribution. In the case of the following dataset I saw a right skewed distribution. Following this idea, a limitation of $ 800 was applied to the dataset. With is, the distribuiton slightly closer to a normal one as you can visualize in the histograms below. I used log price for this purpose.


```{r, echo = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXPLORATORY DATA ANALYSIS ------------------------ #
###### Price ######

# Take log of price
data <- data %>%
  mutate(ln_price = log(price_day))

# Remove extreme values
data <- data %>%
  filter(price_day < 800)

# Price Distribution
price_hist <- ggplot(data, aes( x = price_day)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price")


ln_price_hist <- ggplot(data, aes( x = ln_price)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price (log)")

price_hist_grid <- ggarrange(
  price_hist,
  ln_price_hist,
  nrow = 1
)

price_hist_grid 
```


```{r, include = FALSE, message=FALSE, warning=FALSE}
###### Accommodates ######

# Squares and further values to create for accommodation
data <- data %>%
  mutate(n_accommodates2=n_accommodates^2, ln_accommodates=log(n_accommodates))

price_plot <- ggplot(data = data, aes(x=ln_accommodates, y=ln_price)) +
  geom_point(size=1, colour= "grey", shape=16)+
  labs(x="Number of people accomodated",y="Price")+
  geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  theme_bw()
price_plot

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
###### Beds ######

# Take logs of beds
data <- data %>%
  mutate(ln_beds = log(n_beds + 1))

# Plot a non parametric regression plot
beds_plot <- ggplot(data = data, aes(x= ln_beds, y=ln_price)) +
  geom_point(size=1, colour= "grey", shape=16)+
  labs(x="ln(Number of people accomodated)",y="ln(Price")+
  geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  theme_bw()
beds_plot
```

```{r, include = FALSE, message=FALSE, warning=FALSE}
###### Bathrooms ######
table(data$n_bathrooms)

ggplot(data, aes(x = n_bathrooms)) +
  geom_histogram(fill = "#99d8c9", color = "#2ca25f") +
  ylab("") +
  xlab("No. of bathrooms") +
  theme_bw()

# Pool accommodations with 0,1,2,5 bathrooms
data <- data %>%
  mutate(f_bathroom = cut(n_bathrooms, c(0,1,2,5), labels=c(0,1,2), right = F) )

###### Reviews ######
data <- data %>%
  mutate(ln_number_of_reviews = log(n_number_of_reviews+1))

reviews_plot <- ggplot(data, aes(n_number_of_reviews)) +
  geom_histogram(binwidth = 5, fill = "#99d8c9", color = "#2ca25f", alpha = 0.6, size = 0.25) +
  ylab("") +
  xlab("Number of reviews") +
  theme_bw()
reviews_plot

# Pool number of reviews to 3 categories: none, 1-30 and > 30
data <- data %>%
  mutate(f_number_of_reviews = cut(n_number_of_reviews, c(0,1,30,max(data$n_number_of_reviews)), labels=c(0,1,2), right = F))

ln_reviews_plot <- ggplot(data, aes(ln_number_of_reviews)) +
  geom_histogram(binwidth = 0.5, fill = "#99d8c9", color = "#2ca25f", alpha = 0.6, size = 0.25) +
  ylab("") +
  xlab("Log number of reviews") +
  theme_bw()

reviews_hist_grid <- ggarrange(
  reviews_plot,
  ln_reviews_plot,
  nrow = 1
)

###### Minimum Nights ######
# Pool and categorize the number of minimum nights: 1,2,3, 3+
data <- data %>%
  mutate(f_minimum_nights= cut(n_minimum_nights, c(1,2,3,max(data$n_minimum_nights)), labels=c(1,2,3), right = F))
```
```{r, echo = FALSE, message = FALSE, warning = FALSE}
###### Property Type ######
property_type_box <- ggplot(data, aes(x = f_property_type, y = price_day)) +
  stat_boxplot(aes(group = f_property_type), geom = "errorbar", width = 0.3,
               na.rm=T) +
  geom_boxplot(aes(group = f_property_type),
               size = 0.5, width = 0.6,  fill = c("cyan3","#2ca25f", "green", "blue"),alpha = 0.3, na.rm=T, outlier.shape = NA) +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,300), breaks = seq(0,300,100)) +
  labs(x = "Property type",y = "Price")+
  theme_bw()

prop_with_accomm_box <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                        fill = f_property_type, color= f_property_type)) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8, ) +
 stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodates (Persons)",y = "Price") +
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 400), breaks = seq(0,400, 50)) +
  theme_bw() + theme(legend.position = c(0.26,0.88)) + theme(legend.title = element_blank())

price_prop_type <- ggarrange(
  property_type_box,
  prop_with_accomm_box,
  nrow = 1)
annotate_figure(price_prop_type,bottom = text_grob("Note: Box plots for price, graphs dont show extreme values
                                                   Data source: Inside Airbnb dataset, Amsterdam, Sep 2021", color = "black",vjust = 0.5,hjust = 1.05, x = 1, face = "italic", size = 10))
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
###### accommodation capacity & price ######

accom_hist <- ggplot(data, aes(x = n_accommodates)) + geom_histogram(fill = "#2ca25f")+ theme_bw() +
  labs(x = "Accommodation Capacity")

accom_point <- ggplot(data = data, aes(x=n_accommodates, y=price_day)) +
  geom_point(size=1, colour= "grey", shape=16)+
  labs(x="Number of people accomodated",y="Price")+
  geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  theme_bw() +
    labs(x = "Accommodation Capacity")

accom_combine <- ggarrange(
  accom_hist,
  accom_point,
  nrow = 1)
accom_combine
```
```{r, include = FALSE, warning = FALSE, message = FALSE}
# Change Infinite values with NaNs
for (j in 1:ncol(data) ) data.table::set(data, which(is.infinite(data[[j]])), j, NA)

# change characters to factors
data <- data %>%
  mutate_if(is.character, factor)
```

```{r, echo=FALSE, message=FALSE, error=FALSE}
accommodates_price <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                       fill = factor(d_instant_bookable))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8, fill = "#2ca25f") +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodation Capacity",y = "Price" , fill = "Instant_Bookable")+
  theme_bw() +
  theme(legend.position = "bottom")
accommodates_price
```
## PREDICTIONS

