---
title: "Amsterdam Airbnb Price Prediction"
author: "Nawal Zehra Hasan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r libraries, include = FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
# Import libraries
library(rmdformats)
library(kableExtra)
library(tidyverse)
library(rmdformats)
library(stringr) 
library(rstatix)
library(scales)
library(Metrics)
library(caret)
library(ggpubr)
library(cowplot)
library(dplyr)
library(skimr)
library(modelsummary)
library(grid)
#install.packages("glmnet")
library(glmnet)
library(Hmisc)
library(ranger)
library(RColorBrewer)
library(pdp)
# install.packages("gbm")
library(gbm)
# install.packages("rattle")
library(rattle)
# colours for charts
colours <- c(brewer.pal( 3, "Set2" )[1], brewer.pal( 3, "Set2" )[2], brewer.pal( 3, "Set2" )[3])
```

```{r Import data, message=FALSE, warning=FALSE, include=FALSE, cache=TRUE}
# Import data
data <- read_csv("https://raw.githubusercontent.com/nawalhasan/DA3/main/Assignment2/data/amsterdam_raw.csv")
```
# EXECUTIVE SUMMARY

The idea of the following report is to provide statistical models capable of predicting the rental prices in Amsterdam of apartments accommodating between 2 to 6 persons. With a project like these Airbnbs rentals can estimate at what price should they rent out their apartments and what the profit margins will be at that given price. Using 4 machine learning algorithms; OLS Linear regression, LASSO, and Random Forest & CART I was able to perform this analysis. 

## THE DATASET

To accomplish the given task, I have used the Airbnb data for Amsterdam for the date 7th September 2021 that is available on [Inside Airbnb](www.insideairbnb.com) website. The dataset has a single table that includes 16116 observations.

The target variable given is price per night in USD. The predictors consist of several variables related to property reviews, neighborhood, amenities offered etc. The data set had several variables that would reflect the quality of an Airbnb rental. This shows that data collection was thorough and detailed.

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- FIRST LOOK INTO DATA ------------------------ #
glimpse(data)
skim(data)

# Filtering for apartments which accommodate 2-6 persons
data <- data %>%  filter(between(accommodates, 2, 6))

# summary statistics for the data
data %>%
  group_by(accommodates) %>%
  dplyr::summarize(mean_price = mean(price, na.rm=TRUE))
Hmisc::describe(data$price)

# Check to see if there are any duplicates
duplicated(names(data))

# Drop unused columns
data <- data[grep("^host", colnames(data), invert = TRUE)]
data <- data %>% dplyr::select(-contains("maximum"))
data <- data[grep("^calculated", colnames(data), invert = TRUE)]
data <- data %>% select(-c("listing_url","scrape_id","last_scraped","name","description","neighborhood_overview","picture_url", "neighbourhood_group_cleansed","bathrooms","minimum_minimum_nights","minimum_nights_avg_ntm","calendar_updated","calendar_last_scraped","number_of_reviews_ltm","number_of_reviews_l30d","license","reviews_per_month","availability_30","availability_60","availability_90","availability_365","neighbourhood","has_availability", "first_review", "last_review"))

## Create Numerical variables
data <- data %>%
  mutate(
    price_day = as.numeric(gsub("[^0-9.]", "", price)))

```


```{r, include = FALSE, message=FALSE, warning=FALSE }
# --------------------------- DEAL WITH MISSING VALUES ------------------------ #

# A) drop if no target value
data <- data %>%
  drop_na(price)

# B) Impute if few
data <- data %>%
  mutate(
    beds = ifelse(is.na(beds), accommodates, beds), #assume beds=accomodates
    minimum_nights=ifelse(is.na(minimum_nights),1, minimum_nights),
    number_of_reviews=ifelse(is.na(number_of_reviews),1, number_of_reviews),
    beds=ifelse(is.na(beds),0, beds),
  )

# C) drop columns when many missing: review_scores_value & review_scores_communication
to_drop <- c("review_scores_value", "review_scores_communication")
data <- data %>%
  select(-one_of(to_drop))

to_filter <- sapply(data, function(x) sum(is.na(x)))
to_filter[to_filter > 0]

# D)  Replace missing variables re-reviews with zero, when no review + add flags
data <- data %>%
  mutate(
    flag_review_scores_rating = ifelse(is.na(review_scores_rating),1, 0),
    review_scores_rating    = ifelse(is.na(review_scores_rating),median(review_scores_rating, na.rm = T), review_scores_rating))

# check one flagged variable
datasummary( factor(flag_review_scores_rating) ~ N , data) 

```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- FILTER PROPERTY TYPES ------------------------ #
# Keep if property type is Entire loft, Entire serviced apartment
data <- data %>% 
  filter( property_type %in% c('Entire serviced apartment', 'Entire loft',  'Entire rental unit') )

# rename property_type categories to make them shorter
data <- data %>% mutate( property_type = 
                         ifelse( property_type == 'Entire serviced apartment', 'serviced_apartment',
                         ifelse( property_type == 'Entire loft', 'loft',
                         ifelse( property_type == 'Entire rental unit', 'rental unit',"."))))

# convert property_type to factor
data <- data %>%
  mutate(f_property_type = factor(property_type))

# convert neighborhood_cleansed to factor
data <- data %>% 
  mutate( f_neighbourhood = factor(neighbourhood_cleansed))

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT NUMBER OF BATHROOMS ------------------------ #

data <- data  %>% 
  mutate(bathrooms = as.numeric(gsub("[a-zA-Z ]", "", data$bathrooms_text)))
data$bathrooms_text <- NULL

# add new numeric columns from certain columns
numericals <- c("accommodates", "bathrooms" , "minimum_nights","beds" ,"review_scores_rating","number_of_reviews", "bedrooms")

data <- data %>%
  mutate_at(vars(numericals), funs("n"=as.numeric))

# rename columns so they start with n_ as opposed to end with _n
nnames <- data %>%
  select(ends_with("_n")) %>%
  names()
nnames_i <- match(nnames, colnames(data))
colnames(data)[nnames_i] <- paste0("n_", numericals)
```

```{r,include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT AMENITIES ------------------------ #

# remove unnecessary signs and convert to list
data$amenities <- tolower( data$amenities )
data$amenities <- gsub("\\[","", data$amenities)
data$amenities <- gsub("\\]","", data$amenities)
data$amenities <- gsub('\\"',"",data$amenities)
data$amenities <- as.list(strsplit(data$amenities, ","))

# define levels and dummies and append to df
levs <- levels(factor(unlist(data$amenities)))
data <- cbind(data,as.data.frame(do.call(rbind, lapply(lapply(data$amenities, factor, levs), table))))

# function to aggregate several columns of same type/category into one generic binary column
aggregate_columns <- function(word){
  
  # subset columns which contain a specific word and save them to another dataframe, also select 'id' to use for merge later
  new_df <- data %>% select(contains(word),"id")
  
  # go row by row to see if any of the rows have a 1, if it does, populate new column 'col_name' with 1
  new_df$col_name <- apply(new_df[0:ncol(new_df)], 1, function(x) ifelse(any(x == 1), '1', '0'))
  
  # save new column and id column to another dataframe, this new dataframe is used to merge with original dataframe
  new_df_merge <- new_df %>% select(id,col_name)
  
  # merge original dataframe and new_df_merge by 'id'
  data <- merge(data,new_df_merge,by = "id", all = FALSE)
  
  # remove the new column and 'id' column from the new_df dataframe
  new_df <- new_df %>% select(-c(id,col_name))

  # remove the selected columns from original dataframe since they have already been aggregated into a new column and merged
  data <<- data %>% select(-colnames(new_df))
}

# aggregate columns for a few amenities that could be important for predicting price
aggregate_columns("wifi")
data <- data %>% rename("wifi" = col_name)

aggregate_columns("refrigerator")
data <- data %>% rename("refrigerator" = col_name)

aggregate_columns("air conditioning")
data <- data %>% rename("air_conditioning" = col_name)

aggregate_columns("baby")
data <- data %>% rename("baby" = col_name)

aggregate_columns("beach")
data <- data %>% rename("beach" = col_name)

aggregate_columns("stove")
data <- data %>% rename("stove" = col_name)

aggregate_columns("free parking")
data <- data %>% rename("free_parking" = col_name)

aggregate_columns("office")
data <- data %>% rename("office" = col_name)

aggregate_columns("coffee maker")
data <- data %>% rename("coffee_maker" = col_name)

aggregate_columns("garden")
data <- data %>% rename("garden" = col_name)

aggregate_columns("gym")
data <- data %>% rename("gym" = col_name)

# drop the amenities column because a csv cannot store it since it is a list
data <- data %>% select( -amenities )

# drop amenities that were not used
data <- data[ -c(31:398)]

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- CREATE DUMMY VARIABLES ------------------------ #

data$instant_bookable <- replace(data$instant_bookable,data$instant_bookable == 'TRUE', "1")
data$instant_bookable <- replace(data$instant_bookable,data$instant_bookable == 'FALSE', "0")

# rename dummies
dummies <- c( "instant_bookable", "wifi", "refrigerator", "air_conditioning", "baby", "beach", "stove", "free_parking", "office", "coffee_maker", "garden", "gym")
data <- data %>%
  mutate_at(vars(dummies), funs("d"= (.)))
# rename columns
dnames <- data %>%
  select(ends_with("_d")) %>%
  names()
dnames_i <- match(dnames, colnames(data))
colnames(data)[dnames_i] <- paste0("d_", tolower(gsub("[^[:alnum:]_]", "",dummies)))

# check if price is missing
nrow(data %>% filter( is.na(price_day)))
```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# keep columns if contain d_, n_,f_, p_, usd_ and some others
data <- data %>%
  select(id,price_day,matches("^d_.*|^n_.*|^f_.*"))
amenities_convert<- data %>%
  select(starts_with("d_"),"id") 
amenities_convert <- amenities_convert %>%mutate_if(is.integer,as.numeric)
glimpse(amenities_convert)
data <- data %>%
  select(-starts_with("d_")) 
data <- merge(data,amenities_convert, by = "id")
data <- data %>% mutate(id = as.numeric(id))
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
# Change Infinite values with NaNs
for (j in 1:ncol(data) ) data.table::set(data, which(is.infinite(data[[j]])), j, NA)

# change characters to factors
data <- data %>%
  mutate_if(is.character, factor)
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
write_csv(data,"amsterdam_clean.csv")
```

```{r, echo = FALSE, message=FALSE, warning=FALSE, include=FALSE}
# --------------------------- EXPLORATORY DATA ANALYSIS ------------------------ #
data <- read.csv("amsterdam_clean.csv")

### LABEL ENGINEERING ###
summary(data$price_day)
describe(data$price_day)

# summary table for price
price_stat <- data %>% summarise(
    Variable = 'price_day',
    Mean     = mean( price_day ),
    `5th Percentile` = quantile(price_day, probs = 0.05),
    Median   = median( price_day ),
    `95th Percentile` = quantile(price_day, probs = 0.95),
    Std      = sd( price_day ),
    Min      = min( price_day ),
    Max      = max( price_day ),
    N        = n() )
```

```{r, , echo = FALSE, message=FALSE, warning=FALSE}
# print table
price_stat <- knitr::kable( price_stat, caption = "Descriptive statistics of target variable", digits = 2 ) %>% kable_styling( position = "center", latex_options = 'hold_position' )
price_stat
```

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8}
###### Price ######

# Take log of price
data <- data %>%
  mutate(ln_price = log(price_day))

# prices less than 95th percentile
data <- data %>%
  filter(price_day < 300)

# Price Distribution
price_hist <- ggplot(data, aes( x = price_day)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price")


ln_price_hist <- ggplot(data, aes( x = ln_price)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price (log)")

price_hist_grid <- ggarrange(
  price_hist,
  ln_price_hist,
  nrow = 1
)

price_hist_grid 
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', fig.width=10}
### FEATURE ENGINEERING ###

###### accommodation capacity & price ######

accom_hist <- ggplot(data, aes(x = n_accommodates)) + geom_histogram(fill = "#2ca25f")+ theme_bw() +
  labs(x = "Accommodation Capacity")

accom_point <- ggplot(data = data, aes(x=n_accommodates, y=price_day)) +
  geom_point(size=1, colour= "grey", shape=16)+
  labs(x="Number of people accomodated",y="Price")+
  geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  theme_bw() +
    labs(x = "Accommodation Capacity")

accom_price <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                       fill = factor(d_instant_bookable))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  scale_fill_manual(values=c("#2ca25f", "#99d8c9")) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodates (Persons)",y = "Price" , fill = "Instant_Bookable")+
  theme_bw() +
  theme(legend.position = "bottom")


accom_combine <- ggarrange(
  accom_hist,
  accom_point,
  accom_price,
  nrow = 1)

accom_combine
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center'}
###### Property Type & Price ######
property_type_box <- ggplot(data, aes(x = f_property_type, y = price_day)) +
  stat_boxplot(aes(group = f_property_type), geom = "errorbar", width = 0.3,
               na.rm=T) +
  geom_boxplot(aes(group = f_property_type),
               size = 0.5, width = 0.6,  fill = c("#99d8c9","#2ca25f", "green"),alpha = 0.3, na.rm=T, outlier.shape = NA) +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,300), breaks = seq(0,300,100)) +
  labs(x = "Property type",y = "Price")+
  theme_bw()

prop_with_accomm_box <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                        fill = f_property_type, color= f_property_type)) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8, ) +
 stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodation Capacity",y = "Price") +
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 400), breaks = seq(0,400, 50)) +
  theme_bw() + theme(legend.position = c(0.26,0.88)) + theme(legend.title = element_blank())

price_prop_type <- ggarrange(
  property_type_box,
  prop_with_accomm_box,
  nrow = 1, 
  legend="bottom")
price_prop_type
```

```{r, include = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
# Take logs of beds
data <- data %>%
  mutate(ln_beds = log(n_beds + 1))
ggplot(data, aes(x = n_beds)) + geom_histogram() + theme_bw()
# Plot a non parametric regression plot
beds_plot <- ggplot(data = data, aes(x= ln_beds, y=price)) +
  geom_point(size=1, colour= "cyan3", shape=16)+
  labs(x="ln(Number of people accomodated)",y="ln(Price, Euros")+
  geom_smooth(method="loess", colour= "red", se=FALSE)+
  theme_bw()

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
###### Accommodates ######

# Squares and further values to create for accommodation
data <- data %>%
  mutate(n_accommodates2=n_accommodates^2, ln_accommodates=log(n_accommodates))

#price_plot <- ggplot(data = data, aes(x=ln_accommodates, y=price_day)) +
  #geom_point(size=1, colour= "grey", shape=16)+
  #labs(x="Number of people accomodated",y="Price")+
  #geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  #theme_bw()
#price_plot

```

```{r bathroom chart, include = FALSE, warning = FALSE, message = FALSE}
# Pool accommodations with 0,1,2,5 bathrooms
data <- data %>%
  mutate(f_bathroom = cut(n_bathrooms, c(0,1,2,5), labels=c(0,1,2), right = F) )
```

```{r number of reviews, include = FALSE, warning = FALSE, message = FALSE}
# Pool num of reviews to 3 categories: none, 1-51 and >51
data <- data %>%
  mutate(f_number_of_reviews = cut(n_number_of_reviews, c(0,1,51,max(data$n_number_of_reviews)), labels=c(0,1,2), right = F))
```

```{r minimum_nights, include = FALSE, warning = FALSE, message = FALSE}
# Pool and categorize the number of minimum nights: 1,2,3, 3+
data <- data %>%
  mutate(f_minimum_nights= cut(n_minimum_nights, c(1,2,3,max(data$n_minimum_nights)), labels=c(1,2,3), right = F))
```

```{r Impute NAs, include = FALSE, warning = FALSE, message = FALSE}
# Change Infinite values with NaNs
for (j in 1:ncol(data) ) data.table::set(data, which(is.infinite(data[[j]])), j, NA)
```

```{r Impute missing values, include = FALSE, warning = FALSE, message = FALSE}
# Number of missing values in each column
na_count <- sapply(data, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)

data <- data %>% 
  drop_na(price_day)
# Fill missing values
data <- data %>%
  mutate(
    n_bathrooms =  ifelse(is.na(n_bathrooms), median(n_bathrooms, na.rm = T), n_bathrooms), #assume at least 1 bath
    n_beds = ifelse(is.na(n_beds), n_accommodates, n_beds), #assume n_beds=n_accomodates
    f_bathroom=ifelse(is.na(f_bathroom),1, f_bathroom),
    f_minimum_nights=ifelse(is.na(f_minimum_nights),1, f_minimum_nights),
    f_number_of_reviews=ifelse(is.na(f_number_of_reviews),1, f_number_of_reviews),
    ln_beds=ifelse(is.na(ln_beds),0, ln_beds),
    n_bedrooms=ifelse(is.na(n_bedrooms),1, n_bedrooms)
  )
data <- data %>%
  mutate(
    flag_review_scores_rating=ifelse(is.na(n_review_scores_rating),1, 0),
    n_review_scores_rating =  ifelse(is.na(n_review_scores_rating), median(n_review_scores_rating, na.rm = T), n_review_scores_rating))
```

```{r data type, include = FALSE, warning = FALSE, message = FALSE}
data <- data %>%
  mutate_if(is.character, factor)
```
