---
title: "Amsterdam Airbnb Price Prediction"
author: "Nawal Zehra Hasan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r libraries, include = FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
# Import libraries
library(rmdformats)
library(kableExtra)
library(tidyverse)
library(rmdformats)
library(stringr) 
library(rstatix)
library(scales)
library(Metrics)
library(caret)
library(ggpubr)
library(cowplot)
library(dplyr)
library(skimr)
library(modelsummary)
library(grid)
#install.packages("glmnet")
library(glmnet)
library(Hmisc)
library(ranger)
library(RColorBrewer)
library(pdp)
# install.packages("gbm")
library(gbm)
# install.packages("rattle")
library(rattle)
# colours for charts
colours <- c(brewer.pal( 3, "Set2" )[1], brewer.pal( 3, "Set2" )[2], brewer.pal( 3, "Set2" )[3])
```

```{r Import data, message=FALSE, warning=FALSE, include=FALSE, cache=TRUE}
# Import data
data <- read_csv("https://raw.githubusercontent.com/nawalhasan/DA3/main/Assignment2/data/amsterdam_raw.csv")
```
# EXECUTIVE SUMMARY

The idea of the following report is to provide statistical models capable of predicting the rental prices in Amsterdam of apartments accommodating between 2 to 6 persons. With a project like these Airbnbs rentals can estimate at what price should they rent out their apartments and what the profit margins will be at that given price. Using 4 machine learning algorithms; OLS Linear regression, LASSO, and Random Forest & CART I was able to perform this analysis. 

## THE DATASET

To accomplish the given task, I have used the Airbnb data for Amsterdam for the date 7th September 2021 that is available on [Inside Airbnb](www.insideairbnb.com) website. The dataset has a single table that includes 16116 observations.

The target variable given is price per night in USD. The predictors consist of several variables related to property reviews, neighborhood, amenities offered etc. The data set had several variables that would reflect the quality of an Airbnb rental. This shows that data collection was thorough and detailed.

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- FIRST LOOK INTO DATA ------------------------ #
glimpse(data)
skim(data)

# Filtering for apartments which accommodate 2-6 persons
data <- data %>%  filter(between(accommodates, 2, 6))

# summary statistics for the data
data %>%
  group_by(accommodates) %>%
  dplyr::summarize(mean_price = mean(price, na.rm=TRUE))
Hmisc::describe(data$price)

# Check to see if there are any duplicates
duplicated(names(data))

# Drop unused columns
data <- data[grep("^host", colnames(data), invert = TRUE)]
data <- data %>% dplyr::select(-contains("maximum"))
data <- data[grep("^calculated", colnames(data), invert = TRUE)]
data <- data %>% select(-c("listing_url","scrape_id","last_scraped","name","description","neighborhood_overview","picture_url", "neighbourhood_group_cleansed","bathrooms","minimum_minimum_nights","minimum_nights_avg_ntm","calendar_updated","calendar_last_scraped","number_of_reviews_ltm","number_of_reviews_l30d","license","reviews_per_month","availability_30","availability_60","availability_90","availability_365","neighbourhood","has_availability", "first_review", "last_review"))

## Create Numerical variables
data <- data %>%
  mutate(
    price_day = as.numeric(gsub("[^0-9.]", "", price)))

```


```{r, include = FALSE, message=FALSE, warning=FALSE }
# --------------------------- DEAL WITH MISSING VALUES ------------------------ #

# A) drop if no target value
data <- data %>%
  drop_na(price)

# B) Impute if few
data <- data %>%
  mutate(
    beds = ifelse(is.na(beds), accommodates, beds), #assume beds=accomodates
    minimum_nights=ifelse(is.na(minimum_nights),1, minimum_nights),
    number_of_reviews=ifelse(is.na(number_of_reviews),1, number_of_reviews),
    beds=ifelse(is.na(beds),0, beds),
  )

# C) drop columns when many missing: review_scores_value & review_scores_communication
to_drop <- c("review_scores_value", "review_scores_communication")
data <- data %>%
  select(-one_of(to_drop))

to_filter <- sapply(data, function(x) sum(is.na(x)))
to_filter[to_filter > 0]

# D)  Replace missing variables re-reviews with zero, when no review + add flags
data <- data %>%
  mutate(
    flag_review_scores_rating = ifelse(is.na(review_scores_rating),1, 0),
    review_scores_rating    = ifelse(is.na(review_scores_rating),median(review_scores_rating, na.rm = T), review_scores_rating))

# check one flagged variable
datasummary( factor(flag_review_scores_rating) ~ N , data) 

```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- FILTER PROPERTY TYPES ------------------------ #
# Keep if property type is Entire loft, Entire serviced apartment
data <- data %>% 
  filter( property_type %in% c('Entire serviced apartment', 'Entire loft',  'Entire rental unit') )

# rename property_type categories to make them shorter
data <- data %>% mutate( property_type = 
                         ifelse( property_type == 'Entire serviced apartment', 'serviced_apartment',
                         ifelse( property_type == 'Entire loft', 'loft',
                         ifelse( property_type == 'Entire rental unit', 'rental unit',"."))))

# convert property_type to factor
data <- data %>%
  mutate(f_property_type = factor(property_type))

# convert neighborhood_cleansed to factor
data <- data %>% 
  mutate( f_neighbourhood = factor(neighbourhood_cleansed))

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT NUMBER OF BATHROOMS ------------------------ #

data <- data  %>% 
  mutate(bathrooms = as.numeric(gsub("[a-zA-Z ]", "", data$bathrooms_text)))
data$bathrooms_text <- NULL

# add new numeric columns from certain columns
numericals <- c("accommodates", "bathrooms" , "minimum_nights","beds" ,"review_scores_rating","number_of_reviews", "bedrooms")

data <- data %>%
  mutate_at(vars(numericals), funs("n"=as.numeric))

# rename columns so they start with n_ as opposed to end with _n
nnames <- data %>%
  select(ends_with("_n")) %>%
  names()
nnames_i <- match(nnames, colnames(data))
colnames(data)[nnames_i] <- paste0("n_", numericals)
```

```{r,include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- EXTRACT AMENITIES ------------------------ #

# remove unnecessary signs and convert to list
data$amenities <- tolower( data$amenities )
data$amenities <- gsub("\\[","", data$amenities)
data$amenities <- gsub("\\]","", data$amenities)
data$amenities <- gsub('\\"',"",data$amenities)
data$amenities <- as.list(strsplit(data$amenities, ","))

# define levels and dummies and append to df
levs <- levels(factor(unlist(data$amenities)))
data <- cbind(data,as.data.frame(do.call(rbind, lapply(lapply(data$amenities, factor, levs), table))))

# function to aggregate several columns of same type/category into one generic binary column
aggregate_columns <- function(word){
  
  # subset columns which contain a specific word and save them to another dataframe, also select 'id' to use for merge later
  new_df <- data %>% select(contains(word),"id")
  
  # go row by row to see if any of the rows have a 1, if it does, populate new column 'col_name' with 1
  new_df$col_name <- apply(new_df[0:ncol(new_df)], 1, function(x) ifelse(any(x == 1), '1', '0'))
  
  # save new column and id column to another dataframe, this new dataframe is used to merge with original dataframe
  new_df_merge <- new_df %>% select(id,col_name)
  
  # merge original dataframe and new_df_merge by 'id'
  data <- merge(data,new_df_merge,by = "id", all = FALSE)
  
  # remove the new column and 'id' column from the new_df dataframe
  new_df <- new_df %>% select(-c(id,col_name))

  # remove the selected columns from original dataframe since they have already been aggregated into a new column and merged
  data <<- data %>% select(-colnames(new_df))
}

# aggregate columns for a few amenities that could be important for predicting price
aggregate_columns("wifi")
data <- data %>% rename("wifi" = col_name)

aggregate_columns("refrigerator")
data <- data %>% rename("refrigerator" = col_name)

aggregate_columns("air conditioning")
data <- data %>% rename("air_conditioning" = col_name)

aggregate_columns("baby")
data <- data %>% rename("baby" = col_name)

aggregate_columns("beach")
data <- data %>% rename("beach" = col_name)

aggregate_columns("stove")
data <- data %>% rename("stove" = col_name)

aggregate_columns("free parking")
data <- data %>% rename("free_parking" = col_name)

aggregate_columns("office")
data <- data %>% rename("office" = col_name)

aggregate_columns("coffee maker")
data <- data %>% rename("coffee_maker" = col_name)

aggregate_columns("garden")
data <- data %>% rename("garden" = col_name)

aggregate_columns("gym")
data <- data %>% rename("gym" = col_name)

# drop the amenities column because a csv cannot store it since it is a list
data <- data %>% select( -amenities )

# drop amenities that were not used
data <- data[ -c(31:398)]

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
# --------------------------- CREATE DUMMY VARIABLES ------------------------ #

data$instant_bookable <- replace(data$instant_bookable,data$instant_bookable == 'TRUE', "1")
data$instant_bookable <- replace(data$instant_bookable,data$instant_bookable == 'FALSE', "0")

# rename dummies
dummies <- c( "instant_bookable", "wifi", "refrigerator", "air_conditioning", "baby", "beach", "stove", "free_parking", "office", "coffee_maker", "garden", "gym")
data <- data %>%
  mutate_at(vars(dummies), funs("d"= (.)))
# rename columns
dnames <- data %>%
  select(ends_with("_d")) %>%
  names()
dnames_i <- match(dnames, colnames(data))
colnames(data)[dnames_i] <- paste0("d_", tolower(gsub("[^[:alnum:]_]", "",dummies)))

# check if price is missing
nrow(data %>% filter( is.na(price_day)))
```


```{r, include = FALSE, message=FALSE, warning=FALSE}
# keep columns if contain d_, n_,f_, p_, usd_ and some others
data <- data %>%
  select(id,price_day,matches("^d_.*|^n_.*|^f_.*"))
amenities_convert<- data %>%
  select(starts_with("d_"),"id") 
amenities_convert <- amenities_convert %>%mutate_if(is.integer,as.numeric)
glimpse(amenities_convert)
data <- data %>%
  select(-starts_with("d_")) 
data <- merge(data,amenities_convert, by = "id")
data <- data %>% mutate(id = as.numeric(id))
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
# Change Infinite values with NaNs
for (j in 1:ncol(data) ) data.table::set(data, which(is.infinite(data[[j]])), j, NA)

# change characters to factors
data <- data %>%
  mutate_if(is.character, factor)
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
write_csv(data,"amsterdam_clean.csv")
```

```{r, echo = FALSE, message=FALSE, warning=FALSE, include=FALSE}
# --------------------------- EXPLORATORY DATA ANALYSIS ------------------------ #
data <- read.csv("amsterdam_clean.csv")

### LABEL ENGINEERING ###
summary(data$price_day)
describe(data$price_day)

# summary table for price
price_stat <- data %>% summarise(
    Variable = 'price_day',
    Mean     = mean( price_day ),
    `5th Percentile` = quantile(price_day, probs = 0.05),
    Median   = median( price_day ),
    `95th Percentile` = quantile(price_day, probs = 0.95),
    Std      = sd( price_day ),
    Min      = min( price_day ),
    Max      = max( price_day ),
    N        = n() )
```

```{r, , echo = FALSE, message=FALSE, warning=FALSE}
# print table
price_stat <- knitr::kable( price_stat, caption = "Descriptive statistics of target variable", digits = 2 ) %>% kable_styling( position = "center", latex_options = 'hold_position' )
price_stat
```

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=8}
###### Price ######

# Take log of price
data <- data %>%
  mutate(ln_price = log(price_day))

# prices less than 95th percentile
data <- data %>%
  filter(price_day < 300)

# Price Distribution
price_hist <- ggplot(data, aes( x = price_day)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price")


ln_price_hist <- ggplot(data, aes( x = ln_price)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),fill = "#99d8c9", color = "#2ca25f") +
  theme_bw() +
  scale_y_continuous(labels = label_percent()) +
  ylab("Percent") + 
  xlab("Price (log)")

price_hist_grid <- ggarrange(
  price_hist,
  ln_price_hist,
  nrow = 1
)

price_hist_grid 
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', fig.width=10}
### FEATURE ENGINEERING ###

###### accommodation capacity & price ######

accom_hist <- ggplot(data, aes(x = n_accommodates)) + geom_histogram(fill = "#2ca25f")+ theme_bw() +
  labs(x = "Accommodation Capacity")

accom_point <- ggplot(data = data, aes(x=n_accommodates, y=price_day)) +
  geom_point(size=1, colour= "grey", shape=16)+
  labs(x="Number of people accomodated",y="Price")+
  geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  theme_bw() +
    labs(x = "Accommodation Capacity")

accom_price <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                       fill = factor(d_instant_bookable))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  scale_fill_manual(values=c("#2ca25f", "#99d8c9")) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodates (Persons)",y = "Price" , fill = "Instant_Bookable")+
  theme_bw() +
  theme(legend.position = "bottom")


accom_combine <- ggarrange(
  accom_hist,
  accom_point,
  accom_price,
  nrow = 1)

accom_combine
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center'}
###### Property Type & Price ######
property_type_box <- ggplot(data, aes(x = f_property_type, y = price_day)) +
  stat_boxplot(aes(group = f_property_type), geom = "errorbar", width = 0.3,
               na.rm=T) +
  geom_boxplot(aes(group = f_property_type),
               size = 0.5, width = 0.6,  fill = c("#99d8c9","#2ca25f", "green"),alpha = 0.3, na.rm=T, outlier.shape = NA) +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,300), breaks = seq(0,300,100)) +
  labs(x = "Property type",y = "Price")+
  theme_bw()

prop_with_accomm_box <- ggplot(data, aes(x = factor(n_accommodates), y = price_day,
                        fill = f_property_type, color= f_property_type)) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8, ) +
 stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  labs(x = "Accomodation Capacity",y = "Price") +
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 400), breaks = seq(0,400, 50)) +
  theme_bw() + theme(legend.position = c(0.26,0.88)) + theme(legend.title = element_blank())

price_prop_type <- ggarrange(
  property_type_box,
  prop_with_accomm_box,
  nrow = 1, 
  legend="bottom")
price_prop_type
```

```{r, include = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
# Take logs of beds
data <- data %>%
  mutate(ln_beds = log(n_beds + 1))
ggplot(data, aes(x = n_beds)) + geom_histogram() + theme_bw()
# Plot a non parametric regression plot
beds_plot <- ggplot(data = data, aes(x= ln_beds, y=price)) +
  geom_point(size=1, colour= "cyan3", shape=16)+
  labs(x="ln(Number of people accomodated)",y="ln(Price, Euros")+
  geom_smooth(method="loess", colour= "red", se=FALSE)+
  theme_bw()

```

```{r, include = FALSE, message=FALSE, warning=FALSE}
###### Accommodates ######

# Squares and further values to create for accommodation
data <- data %>%
  mutate(n_accommodates2=n_accommodates^2, ln_accommodates=log(n_accommodates))

#price_plot <- ggplot(data = data, aes(x=ln_accommodates, y=price_day)) +
  #geom_point(size=1, colour= "grey", shape=16)+
  #labs(x="Number of people accomodated",y="Price")+
  #geom_smooth(method="loess", colour= "#2ca25f", se=FALSE)+
  #theme_bw()
#price_plot

```

```{r bathroom chart, include = FALSE, warning = FALSE, message = FALSE}
# Pool accommodations with 0,1,2,5 bathrooms
data <- data %>%
  mutate(f_bathroom = cut(n_bathrooms, c(0,1,2,5), labels=c(0,1,2), right = F) )
```

```{r number of reviews, include = FALSE, warning = FALSE, message = FALSE}
# Pool num of reviews to 3 categories: none, 1-51 and >51
data <- data %>%
  mutate(f_number_of_reviews = cut(n_number_of_reviews, c(0,1,51,max(data$n_number_of_reviews)), labels=c(0,1,2), right = F))
```

```{r minimum_nights, include = FALSE, warning = FALSE, message = FALSE}
# Pool and categorize the number of minimum nights: 1,2,3, 3+
data <- data %>%
  mutate(f_minimum_nights= cut(n_minimum_nights, c(1,2,3,max(data$n_minimum_nights)), labels=c(1,2,3), right = F))
```

```{r Impute NAs, include = FALSE, warning = FALSE, message = FALSE}
# Change Infinite values with NaNs
for (j in 1:ncol(data) ) data.table::set(data, which(is.infinite(data[[j]])), j, NA)
```

```{r Impute missing values, include = FALSE, warning = FALSE, message = FALSE}
# Number of missing values in each column
na_count <- sapply(data, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)

data <- data %>% 
  drop_na(price_day)
# Fill missing values
data <- data %>%
  mutate(
    n_bathrooms =  ifelse(is.na(n_bathrooms), median(n_bathrooms, na.rm = T), n_bathrooms), #assume at least 1 bath
    n_beds = ifelse(is.na(n_beds), n_accommodates, n_beds), #assume n_beds=n_accomodates
    f_bathroom=ifelse(is.na(f_bathroom),1, f_bathroom),
    f_minimum_nights=ifelse(is.na(f_minimum_nights),1, f_minimum_nights),
    f_number_of_reviews=ifelse(is.na(f_number_of_reviews),1, f_number_of_reviews),
    ln_beds=ifelse(is.na(ln_beds),0, ln_beds),
    n_bedrooms=ifelse(is.na(n_bedrooms),1, n_bedrooms)
  )
data <- data %>%
  mutate(
    flag_review_scores_rating=ifelse(is.na(n_review_scores_rating),1, 0),
    n_review_scores_rating =  ifelse(is.na(n_review_scores_rating), median(n_review_scores_rating, na.rm = T), n_review_scores_rating))
```

```{r data type, include = FALSE, warning = FALSE, message = FALSE}
data <- data %>%
  mutate_if(is.character, factor)
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
# helper function ----------------------------------------------------------
price_diff_by_variables2 <- function(df, factor_var, dummy_var, factor_lab, dummy_lab){
  # Looking for interactions.
  # It is a function it takes 3 arguments: 1) Your dataframe,
  # 2) the factor variable (like room_type)
  # 3)the dummy variable you are interested in (like TV)
  # Process your data frame and make a new dataframe which contains the stats
  factor_var <- as.name(factor_var)
  dummy_var <- as.name(dummy_var)
  stats <- df %>%
    group_by(!!factor_var, !!dummy_var) %>%
    dplyr::summarize(Mean = mean(price_day, na.rm=TRUE),
                     se = sd(price_day)/sqrt(n()))
  stats[,2] <- lapply(stats[,2], factor)
  ggplot(stats, aes_string(colnames(stats)[1], colnames(stats)[3], fill = colnames(stats)[2]))+
    geom_bar(stat='identity', position = position_dodge(width=0.9), alpha=0.8)+
    geom_errorbar(aes(ymin=Mean-(1.96*se),ymax=Mean+(1.96*se)),
                  position=position_dodge(width = 0.9), width = 0.25)+
    scale_color_manual(name=dummy_lab,
                       values=c("#99d8c9", "#2ca25f")) +
    scale_fill_manual(name=dummy_lab,
                      values= c("#99d8c9", "#2ca25f")) +
    ylab('Mean Price')+
    xlab(factor_lab) +
    theme(panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          panel.border=element_blank(),
          axis.line=element_line(),
          legend.position = "top",
          #legend.position = c(0.7, 0.9),
          legend.box = "vertical",
          legend.text = element_text(size = 5),
          legend.title = element_text(size = 5, face = "bold"),
          legend.key.size = unit(x = 0.4, units = "cm")
        )
}
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
p1 <- price_diff_by_variables2(data, "f_property_type", "d_wifi", "Property Type", "Wifi")
p2 <- price_diff_by_variables2(data, "f_property_type", "d_refrigerator", "Property Type", "Regrigerator")
p3 <- price_diff_by_variables2(data, "f_property_type", "d_instant_bookable" , "Property Type", "Instant Bookable") 
p4 <- price_diff_by_variables2(data, "f_property_type", "d_air_conditioning" , "Property Type", "Air Conditioning")
p5 <- price_diff_by_variables2(data, "f_property_type", "d_stove", "Property Type", "Stove")
p6 <- price_diff_by_variables2(data, "f_property_type", "d_baby", "Property Type", "Baby Friendly") 
p7 <- price_diff_by_variables2(data, "f_property_type", "d_beach", "Property Type", "Beach")
p8 <- price_diff_by_variables2(data, "f_property_type", "d_free_parking", "Property Type", "Free Parking") # dont add to interactions
p9 <- price_diff_by_variables2(data, "f_property_type", "d_office", "Property Type", "Office Space")
p10 <- price_diff_by_variables2(data, "f_property_type", "d_coffee_maker", "Property Type", "Coffee Maker")
p11 <- price_diff_by_variables2(data, "f_property_type", "d_garden", "Property Type", "Garden")
p12 <- price_diff_by_variables2(data, "f_property_type", "d_gym", "Property Type", "Gym")
sum_interactions <- plot_grid(p10, p4, p12, p6, nrow=2, ncol=2)
sum_interactions
```

```{r, include = FALSE, cache=TRUE}
# dummies suggested by graphs
X1  <- c("f_property_type*d_instant_bookable", "f_property_type*d_refrigerator", "f_property_type*d_coffee_maker" )
X2 <- c("f_property_type*d_stove", "f_property_type*d_baby", "f_property_type*d_beach", "f_property_type*d_gym")
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
# Assign columns to grouped variables for model equations
n_var <- c("n_accommodates", "n_bedrooms", "n_review_scores_rating")
f_var <- c("f_property_type", "f_minimum_nights", "f_number_of_reviews", "f_bathroom")
poly_var <- c("n_accommodates2", "ln_beds")
# Dummy variables: Extras -> collect all options and create dummies
d_amenities <-  grep("^d_.*", names(data), value = TRUE)

m1 <- "= Number of guests accommodated, number of bedrooms, average review scores, guests accommodated (squared term), property type, minimum nights, number of reviews, number of bathrooms, log number of beds,"
m2 <- "= M1 + all amenities"
m3 <- "= M2 + amenities interactions"
model_variables <- c(m1,m2,m3)
model_names <- c("M1", "M2", "M3")
model_table <- as.data.frame(cbind(model_names, model_variables))
model_headings <- c("Model", "Predictor Variables")
colnames(model_table) <- model_headings
```

```{r model table, echo = FALSE, warning = FALSE, message = FALSE}
model_table %>%
  kbl(caption = "<center><strong>Versions of the Airbnb Apartment Price Prediction Models</strong></center>", escape = FALSE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling( position = "center")
```

```{r model construction,  include = FALSE, warning = FALSE, message = FALSE}
# Create models in levels models: 1-3
model1 <- as.formula(paste("price_day ~ ",paste(c(n_var,poly_var, f_var),collapse = " + ")))
model2 <- as.formula(paste("price_day ~ ",paste(c(n_var,poly_var, f_var, d_amenities),collapse = " + ")))
model3 <- as.formula(paste("price_day ~ ",paste(c(n_var,poly_var, f_var, d_amenities, X1, X2),collapse = " + ")))
```

```{r cross validation lm, include = FALSE, warning = FALSE, message = FALSE}
# ------------------------------- OLS ------------------------------- #
#final check for missing values
to_filter <- sapply(data, function(x) sum(is.na(x)))
to_filter[to_filter > 0]

set.seed(8872)
# Create models in levels models: 1-3
# Create models in levels models: 1-3
train_indices <- as.integer(createDataPartition(data$price_day, p = 0.7, list = FALSE))
data_train <- data[train_indices, ]
data_holdout <- data[-train_indices, ]
# model 1 CV
set.seed(8872)
cv_model1 <- train(model1, 
                   data = data_train, 
                   method = "lm",
                   trControl = trainControl(method = "cv", number = 5)
)
# model 2 CV
set.seed(8872)
cv_model2 <- train(
  model2, 
  data = data_train,
  method = "lm",
  trControl = trainControl(method = "cv", number = 5)
)
# model 3 CV
set.seed(8872)
cv_model3 <- train(
  model3, 
  data = data_train, 
  method = "lm",
  trControl = trainControl(method = "cv", number = 5)
)
cv_mod1_pred <- predict(cv_model1, data_train)
cv_mod2_pred <- predict(cv_model2, data_train)
cv_mod3_pred <- predict(cv_model3, data_train)
# Checking coefficients
cv_model1$finalModel # coefficients
# RMSE fold results for all models
model1_rmse <- as.matrix(round(cv_model1$resample$RMSE,3))
model2_rmse <- as.matrix(round(cv_model2$resample$RMSE,3))
model3_rmse <- as.matrix(round(cv_model3$resample$RMSE,3))
mean_rmse <- c(mean(model1_rmse), mean(model2_rmse),mean(model3_rmse))
model_rmse_table <- as.data.frame(cbind(model1_rmse,model2_rmse, model3_rmse))
colnames(model_rmse_table) <- c("Model 1", "Model 2", "Model 3")
model_rmse_table <- rbind(model_rmse_table,mean_rmse)
rownames(model_rmse_table) <- c("Fold 1", "Fold 2", "Fold 3", "Fold 4", "Fold 5", "Average")
#### Comparing Fit measures
model_list <- c(model1,model2,model3)
BIC <- NULL
nvars <- NULL
r2 <- NULL
for(x in model_list){
  model_work_data <- lm(x,data = data_train)
  BIC <- c(BIC,round(BIC(model_work_data)))
  nvars <- c(nvars, model_work_data$rank -1)
  r2 <- c(r2,summary(model_work_data)$r.squared)
}
# Calculate RMSE for training set
rmse_train <- c(mean(cv_model1$resample$RMSE),mean(cv_model2$resample$RMSE), mean(cv_model3$resample$RMSE))
# Calculate RMSE for testing set
rmse_test <- c(rmse(cv_mod1_pred,data_train$price),rmse(cv_mod2_pred,data_train$price), rmse(cv_mod3_pred,data_train$price))
# Bind all the different model results together
model_results <- as.data.frame(cbind(nvars,r2,BIC,rmse_train,rmse_test))
# Convert all numeric columns to numeric data type
model_results <- model_results %>% 
  mutate_if(is.character, numeric)
# Round all numeric columns to 2 digits if applicable
model_results <- model_results %>% 
  mutate_if(is.numeric, round, digits = 2)
# Add model names to the model results table
model_names <- c("Model 1","Model 2","Model 3")
model_results <- cbind(model_names,model_results)
# Create column name list for model results table
column_names <- c("Model", "N predictors", "R-squared", "BIC", "Training RMSE","Test RMSE")
colnames(model_results) <- column_names
#### Holdout set predictions
cv_holdout_pred <- predict(cv_model3, data_holdout)
holdout_rmse <- mean(cv_model3$resample$RMSE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
model_results %>%
  kbl(caption = "<center><strong>Comparisng Model Fit measures</strong></center>", escape = FALSE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling( position = "center")
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
# --------------------------- lASSO ---------------------------- #
model4 <- as.formula(paste("price_day ~ ",paste(c(n_var,poly_var, f_var, d_amenities, X1, X2),collapse = " + ")))
# Set lasso tuning parameters
train_control <- trainControl(
  method = "cv",
  number = 5)
tune_grid <- expand.grid("alpha" = c(1), "lambda" = seq(0.05, 1, by = 0.05))
set.seed(8872)
lasso_model <- caret::train(model4,
                            data = data_train,
                            method = "glmnet",
                            preProcess = c("center", "scale"),
                            trControl = train_control,
                            tuneGrid = tune_grid,
                            na.action=na.exclude)
print(lasso_model$bestTune$lambda) #0.25 RMSE
lasso_coeffs <- coef(lasso_model$finalModel, lasso_model$bestTune$lambda) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "variable") 
print(lasso_coeffs)
# Evaluate model. CV error:
lasso_cv_rmse <- lasso_model$results %>%
  filter(lambda == lasso_model$bestTune$lambda) %>%
  dplyr::select(RMSE)
print(lasso_cv_rmse[1, 1]) #RMSE 44.71
plot(lasso_model)
```

```{r,  lasso,  include = FALSE, warning = FALSE, message = FALSE}
lasso_coeffs %>% kbl(caption = "<center><strong>Lasso Model Coefficients</strong></center>", escape = FALSE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling( position = "center")
```

```{r, echo = FALSE, warning = FALSE, MESSAGE = FALSE}
m1_rf <- "= guests accommodated, number of beds, number of bedrooms, average review scores,"
m2_rf <- "= M1 + f_property_type, f_minimum_nights,f_number_of_reviews, f_bathroom,"
m3_rf <- "= M3 + all amenities columns"
model_variables_rf <- c(m1_rf,m2_rf,m3_rf)
model_names_rf <- c("M1", "M2", "M3")
model_table_rf <- as.data.frame(cbind(model_names_rf, model_variables_rf))
model_headings_rf <- c("Model", "Predictor Variables")
colnames(model_table_rf) <- model_headings_rf
model_table_rf %>%
  kbl(caption = "<center><strong>Versions of the Airbnb Apartment Price Prediction Models for Random Forest</strong></center>", escape = FALSE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling( position = "center")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, include=FALSE}
# ----------------------------- RANDOM FOREST ---------------------------- #
predictors_1 <- c(n_var)
predictors_2 <- c(n_var, f_var)
predictors_3 <- c(n_var, f_var,d_amenities)

train_control <- trainControl(method = "cv",
                              number = 5,
                              verboseIter = FALSE)
# set tuning
tune_grid <- expand.grid(
  .mtry = c(2),
  .splitrule = "variance",
  .min.node.size = c(50)
)

# MODEL 1
# simpler model for model - using random forest
set.seed(8872)
system.time({
rf_model_1 <- train(
  formula(paste0("price_day ~", paste0(predictors_1, collapse = " + "))),
  data = data_train,
  method = "ranger",
  trControl = train_control,
  tuneGrid = tune_grid,
  importance = "impurity"
)
})
rf_model_1

## MODEL 2
set.seed(8872)
system.time({
rf_model_2 <- train(
  formula(paste0("price_day ~", paste0(predictors_2, collapse = " + "))),
  data = data_train,
  method = "ranger",
  trControl = train_control,
  tuneGrid = tune_grid,
  importance = "impurity"
)
})
rf_model_2

# MODEL 3
# simpler model for model - using random forest
set.seed(8872)
system.time({
rf_model_3 <- train(
  formula(paste0("price_day ~", paste0(predictors_3, collapse = " + "))),
  data = data_train,
  method = "ranger",
  trControl = train_control,
  tuneGrid = tune_grid,
  importance = "impurity"
)
})
rf_model_3

results <- resamples(
  list(
    model_1  = rf_model_1,
    model_2  = rf_model_2,
    model_3  = rf_model_3
  )
)
summary(results)

```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# RMSE fold results for all models
model1_rf_rmse <- as.matrix(round(results$values$`model_1~RMSE`,3))
model2_rf_rmse <- as.matrix(round(results$values$`model_2~RMSE`,3))
model3_rf_rmse <- as.matrix(round(results$values$`model_3~RMSE`,3))
mean_rf_rmse <- c(mean(model1_rf_rmse), mean(model2_rf_rmse), mean(model3_rf_rmse))
model_rf_rmse_table <- as.data.frame(cbind(model1_rf_rmse,model2_rf_rmse,model3_rf_rmse))
colnames(model_rf_rmse_table) <- c("Model 1", "Model 2", "Model 3")
model_rf_rmse_table <- rbind(model_rf_rmse_table,mean_rf_rmse)
rownames(model_rf_rmse_table) <- c("Fold 1", "Fold 2", "Fold 3", "Fold 4", "Fold 5", "Average")
model_rf_rmse_table %>% kbl(caption = "<center><strong>RMSE fold results for all models</strong></center>", escape = FALSE) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling( position = "center")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', fig.width=10}
# ----------------------------- VARIABLE IMPORTANCE ---------------------------- #
rf_model_3_var_imp <- ranger::importance(rf_model_3$finalModel)/1000
rf_model_3_var_imp_df <-
  data.frame(varname = names(rf_model_3_var_imp),imp = rf_model_3_var_imp) %>%
  #mutate(varname = gsub("f_neighbourhood_cleansed", "Borough:", varname) ) %>%
  #mutate(varname = gsub("f_room_type", "Room type:", varname) ) %>%
  arrange(desc(imp)) %>%
  mutate(imp_percentage = imp/sum(imp))

# top 10 most important variables
var_imp_a <- ggplot(rf_model_3_var_imp_df[1:10,], aes(x=reorder(varname, imp), y=imp_percentage)) +
  geom_point(color='#2ca25f', size=1) +
  geom_segment(aes(x=varname,xend=varname,y=0,yend=imp_percentage), color='#2ca25f', size=0.75) +
  ylab("Importance (Percent)") +
  xlab("Variable Name") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw()

# grouped variable importance
varnames <- rf_model_3$finalModel$xNames
f_bathroom_varnames <-  grep("f_bathroom",varnames, value = TRUE)
f_property_type_varnames <- grep("f_property_type",varnames, value = TRUE)
f_minimum_nights_varnames <- grep("f_minimum_nights",varnames, value = TRUE)
f_number_of_reviews_varnames <- grep("f_number_of_reviews",varnames, value = TRUE)

groups <- list(f_bathroom= f_bathroom_varnames,
               f_minimum_nights = f_minimum_nights_varnames,
               f_property_type = f_property_type_varnames,
               f_number_of_reviews = f_number_of_reviews_varnames,
               n_accommodates = "n_accommodates")

# Need a function to calculate grouped varimp
group.importance <- function(rf.obj, groups) {
  var.imp <- as.matrix(sapply(groups, function(g) {
    sum(ranger::importance(rf.obj)[g], na.rm = TRUE)
  }))
  colnames(var.imp) <- "MeanDecreaseGini"
  return(var.imp)
}

rf_model_3_var_imp_grouped <- group.importance(rf_model_3$finalModel, groups)
rf_model_3_var_imp_grouped_df <- data.frame(varname = rownames(rf_model_3_var_imp_grouped),
                                            imp = rf_model_3_var_imp_grouped[,1])  %>%
                                      mutate(imp_percentage = imp/sum(imp))

var_imp_b <- ggplot(rf_model_3_var_imp_grouped_df, aes(x=reorder(varname, imp), y=imp_percentage)) +
  geom_point(color='#2ca25f', size=1) +
  geom_segment(aes(x=varname,xend=varname,y=0,yend=imp_percentage), color='#2ca25f', size=0.7) +
  ylab("Importance (Percent)") +   xlab("Variable Name") +
  coord_flip() +
  # expand=c(0,0),
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw()

ggarrange(var_imp_a,var_imp_b, nrow = 1)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.width=12, fig.align='center'}
# ----------------------------- PARTIAL DEPENDENCIES PLOT ---------------------------- #
# 1) Number of accommodates
pdp_n_acc <- pdp::partial(rf_model_3, pred.var = "n_accommodates", 
                          pred.grid = distinct_(data_holdout, "n_accommodates"), 
                          train = data_train)

pdp1 <- pdp_n_acc %>%
  autoplot( ) +
  geom_point(color='#2ca25f', size=2) +
  geom_line(color='#2ca25f', size=1) +
  ylab("Predicted price") +
  xlab("Accommodates (persons)") +
  scale_x_continuous(limit=c(1,7), breaks=seq(1,7,1))+
theme_bw()


# 2) Property type
pdp_n_propertytype <- pdp::partial(rf_model_3, pred.var = "f_property_type", 
                               pred.grid = distinct_(data_holdout, "f_property_type"), 
                               train = data_train)
pdp2 <- pdp_n_propertytype %>%
  autoplot( ) +
  geom_point(color='#2ca25f', size=4) +
  ylab("Predicted price") +
  xlab("Property type") +
  scale_y_continuous(limits=c(120,180), breaks=seq(120,180, by=20)) +
  theme_bw()

# 3) bedrooms
pdp_n_bedrooms <- pdp::partial(rf_model_3, pred.var = "n_bedrooms", 
                               pred.grid = distinct_(data_holdout, "n_bedrooms"), 
                               train = data_train)
pdp3 <- pdp_n_bedrooms %>%
  autoplot( ) +
  geom_point(color='#2ca25f', size=4) +
  ylab("Predicted price") +
  xlab("No. of bedrooms") +
  scale_y_continuous(limits=c(120,180), breaks=seq(120,180, by=20)) +
  theme_bw()

ggarrange(pdp1, pdp2, pdp3,nrow = 1)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, include=FALSE}
# ----------------------------- CART ---------------------------- #
## CART
set.seed(8872)
system.time({
cart_model <- train(
  formula(paste0("price_day ~", paste0(predictors_2, collapse = " + "))),
  data = data_train,
  method = "rpart",
  tuneLength = 10,
  trControl = train_control
)
})
cart_model
# Showing an alternative for plotting a tree
fancyRpartPlot(cart_model$finalModel, sub = "")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, include=FALSE}
#FOR PRACTISE ONLY
# GBM
#gbm_grid <-  expand.grid(interaction.depth = 5, # complexity of the tree
                         #n.trees = 250, # number of iterations, i.e. trees
                         #shrinkage = 0.1, # learning rate: how quickly the algorithm adapts
                         #n.minobsinnode = 20 # the minimum number of training set samples in a node to commence splitting
#)


#set.seed(8872)
#system.time({
  #gbm_model <- train(formula(paste0("price_day ~", paste0(predictors_2, collapse = " + "))),
                     #data = data_train,
                     #method = "gbm",
                     #trControl = train_control,
                     #verbose = FALSE,
                     #tuneGrid = gbm_grid)
#})
#gbm_model
#gbm_model$finalModel
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# ----------------------------- MODEL SELECTION ---------------------------- #
# model comparision
final_models <-
  list("OLS Model 2" = cv_model2,
  "LASSO (model with interactions)" = lasso_model,
  "CART" = cart_model,
  "Random forest(with amenities)" = rf_model_3)
results <- resamples(final_models) %>% summary()

# Save output --------------------------------------------------------
# Model selection is carried out on this CV RMSE
result_4 <- imap(final_models, ~{
  round(mean(results$values[[paste0(.y,"~RMSE")]]),3)
}) %>% unlist() %>% as.data.frame() %>%
  rename("CV RMSE" = ".")

result_5 <- map(final_models, ~{
  RMSE(predict(.x, newdata = data_holdout), data_holdout[["price_day"]])
}) %>% unlist() %>% as.data.frame() %>%
  rename("Holdout RMSE" = ".")

final_combined <- cbind(result_4, result_5)
```
```{r, horserace, echo = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
knitr::kable( final_combined, caption = "Model performance comparison", digits = 2 ) %>% kable_styling( position = "center", latex_options = 'hold_position')
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# ----------------------------- SUBSAMPLE PERFORMANCE ---------------------------- #
# Subsample performance: RMSE / mean(y)
data_holdout_w_prediction <- data_holdout %>%
  mutate(predicted_price = predict(rf_model_2, newdata = data_holdout))

######### create nice summary table of heterogeneity
a <- data_holdout_w_prediction %>%
  mutate(is_low_size = ifelse(n_accommodates <= 3, "small apt", "large apt")) %>%
  group_by(is_low_size) %>%
  dplyr::summarise(
    rmse = RMSE(predicted_price, price_day),
    mean_price = mean(price_day),
    rmse_norm = RMSE(predicted_price, price_day) / mean(price_day)
  )

b <- data_holdout_w_prediction %>%
  dplyr::summarise(
    rmse = RMSE(predicted_price, price_day),
    mean_price = mean(price_day),
    rmse_norm = RMSE(predicted_price, price_day) / mean(price_day)
  )

# Save output
colnames(a) <- c("", "RMSE", "Mean price", "RMSE/price")
b <- cbind("All", b)
colnames(b) <- c("", "RMSE", "Mean price", "RMSE/price")

line1 <- c("Apartment size", "", "", "")

result_3 <- rbind(line1, a, b) %>%
  transform(RMSE = as.numeric(RMSE), `Mean price` = as.numeric(`Mean price`),
            `RMSE/price` = as.numeric(`RMSE/price`))

```


